---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6)
```

### Lost transcript analysis

```{r}
library(tidyverse)
library(tidytext)
library(scales)
library(glue)
library(ggimage)
library(jpeg)
library(grid)
```


```{r}
theme_update(plot.title = element_text(size = 18, face = "bold"),
          plot.subtitle = element_text(size = 12, color = "grey30"))
```


Below reads in the data I scraped from lostpedia.fandom.com. The Rmarkdown to scrape episode transcripts can be found in the scrape_episode_transcripts.Rmd file.

```{r}
lost <- read_csv("data/lost_transcripts.csv")

#remove season 0 (Lost: Missing Pieces)
lost <- lost %>%
    filter(season != 0)
```

<br />

### Which Lost characters have the most lines?

```{r}
bg_img <- readJPEG("images/dharma_logo_alpha_10.jpg")

top_n  <- 30

lost %>%
    mutate(character = fct_lump(character, top_n)) %>%
    filter(character != "Other") %>%
    count(character) %>%
    mutate(character = fct_reorder(character, n)) %>%
    ggplot(aes(n, character)) +
    annotation_custom(rasterGrob(bg_img, width = unit(1, "npc"), height = unit(1, "npc")), 
                  -Inf, Inf, -Inf, Inf) +
    geom_col(alpha = 0.85) +
    labs(x = "Lines",
         y = NULL,
         title = "Jack, Sawyer, and Locke have the most lines in LOST",
         subtitle = glue("The top {top_n} characters are shown")) +
    scale_x_continuous(labels = comma_format())
```

<br />

### Which characters have the most lines per season?

```{r}
# https://stackoverflow.com/questions/44688623/adding-custom-images-to-ggplot-facets
# https://lostpedia.fandom.com/wiki/DHARMA_Initiative_stations

lost %>%
    count(season, character) %>%
    group_by(season) %>%
    slice_max(n, n = 20) %>% 
    ungroup() %>%
    mutate(character = if_else(character == "Christian Shephard", "Christian", character)) %>%
    mutate(character = reorder_within(character, n, season)) %>%
    mutate(season = glue("Season {season}")) %>%
    ggplot(aes(n, character, fill = season)) +
    geom_col() +
    facet_wrap(~season, scales = "free") +
    labs(x = "Count",
         y = NULL,
         title = "Jack has the most lines for all seasons except season 5",
         subtitle = "The top 20 characters are shown") +
    scale_y_reordered() +
    theme(legend.position = "none")
```

<br />

### How has the characters with the most lines changed over six seasons?

```{r}
character_counts <- lost %>%
    count(character, sort = TRUE)

top_n_characters <- 9

lost %>%
    semi_join(character_counts %>% head(top_n_characters)) %>%
    count(season, character) %>%
    group_by(season) %>%
    mutate(rank = rank(-n)) %>%
    ungroup() %>%
    mutate(label_s1 = ifelse(season == 1, character, "")) %>%
    mutate(label_s6 = ifelse(season == 6, character, "")) %>%
    mutate(character = fct_reorder(character, rank, .fun = mean)) %>%
    mutate(character = fct_rev(character)) %>% 
    ggplot(aes(season, rank, group = character)) +
    annotation_custom(rasterGrob(bg_img, width = unit(1, "npc"), height = unit(1, "npc")), 
              -Inf, Inf, -Inf, Inf) +
    geom_line(aes(color = character), size = 5, alpha = 0.75) +
    geom_point(aes(color = character), size = 7) +
    geom_point(size = 1.5, color = "white") +
    geom_text(aes(label = label_s1), nudge_x = -0.5) +
    geom_text(aes(label = label_s6), nudge_x = 0.5) +
    scale_y_reverse(breaks = 1:top_n_characters, minor_breaks = 1) +
    scale_x_continuous(breaks = 1:6) +
    labs(x = "Season",
         y = "Rank number of lines",
         color = NULL,
         title = "Jack has the most lines for all seasons except season 5",
         subtitle = "Note: Desmond and Ben have no rank before they arrive in seasons 2 and 3, respectively") +
    scale_color_manual(values = c("#909090", "#7e7e7e", "#6c6c6c", "#5a5a5a", 
                                  "#484848", "#363636", "#242424", "#121212", "#000000")) +
    theme(legend.position = "none")
```

<br />

### Which words are most characteristic of each Lost season?

```{r}
library(tidylo)

lost %>%
    unnest_tokens(word, line) %>%
    count(season, word, sort = TRUE) %>%
    bind_log_odds(season, word, n) %>%
    group_by(season) %>%
    slice_max(log_odds_weighted, n = 20) %>%
    ungroup() %>%
    mutate(word = reorder_within(word, log_odds_weighted, season)) %>%
    mutate(season = glue("Season {season}")) %>%
    ggplot(aes(log_odds_weighted, word, fill = factor(season))) +
    geom_col() +
    facet_wrap(~season, scales = "free_y") +
    scale_y_reordered() +
    theme(legend.position = "none") +
    labs(x = "Weighted log odds",
         y = NULL,
         title = "Some words are very characteristic of certain Lost seasons",
         subtitle = "Notable mentions are 'chopper' in season 4 and 'LaFleur' in season 5")
```

<br />

### Which words are most characteristic of each Lost character?

```{r}
lost %>%
    semi_join(character_counts %>% head(top_n_characters)) %>%
    unnest_tokens(word, line) %>%
    count(character, word, sort = TRUE) %>%
    bind_log_odds(character, word, n) %>%
    group_by(character) %>%
    slice_max(log_odds_weighted, n = 15) %>%
    ungroup() %>%
    mutate(word = reorder_within(word, log_odds_weighted, character)) %>%
    ggplot(aes(log_odds_weighted, word, fill = character)) +
    geom_col() +
    facet_wrap(~character, scales = "free_y") +
    scale_y_reordered() +
    theme(legend.position = "none") +
    labs(x = "Weighted log odds",
         y = NULL,
         title = "Some words are very characteristic of certain Lost characters",
         subtitle = "Notable mentions are 'dude' for Hurley and 'Freckles' for Sawyer")
```

<br />

### Which emotions are most related to which Lost characters?

```{r}
library(textdata)

lost %>%   
    semi_join(character_counts %>% head(9)) %>%
    unnest_tokens(word, line) %>%
    inner_join(lexicon_nrc()) %>%
    add_count(character) %>%
    count(character, sentiment, n, name = "sentiment_n") %>%
    mutate(prop = sentiment_n / n) %>%
    mutate(character = reorder_within(character, prop, sentiment)) %>%
    ggplot(aes(prop, character, fill = sentiment)) +
    geom_col() +
    facet_wrap(~sentiment, scales = "free_y") +
    scale_y_reordered() +
    theme(legend.position = "none") +
    labs(x = "Proportion of words",
         y = NULL,
         title = "Different characters have different core emotions their word choice conveys",
         subtitle = "Notable mentions are anger for Sawyer and positive emotions for Desmond")
```

<br />

### Which Lost characters are most similar in the emotionality of their words?

This analysis compares the emotional profile for each of the top {r top_n_characters} characters to determine which characters are most similar based on their use of language.

```{r}
library(widyr)
    
lost %>%   
    semi_join(character_counts %>% head(top_n_characters)) %>%
    unnest_tokens(word, line) %>%
    inner_join(lexicon_nrc()) %>%
    add_count(character) %>%
    count(character, sentiment, n, name = "sentiment_n") %>%
    mutate(prop = sentiment_n / n) %>%
    pairwise_cor(character, sentiment, prop) %>% 
    mutate(item1 = reorder_within(item1, correlation, item2)) %>%
    ggplot(aes(correlation, item1, fill = item2)) +
    geom_col() +
    facet_wrap(~item2, scales = "free_y") +
    scale_y_reordered() +
    theme(legend.position = "none") +
    labs(x = "Correlation",
         y = NULL,
         title = "Similarities of characters based on the emotional profile of words they use")
```

<br />
<br />
<br />
<br />
<br />